Staged Abstract Interpreters

### 2019-08-16 paper/30_stageconc.tex
update

The IR nodes are manipulated by the LMS passes and finally generated to
next-stage Scala code. To generate code, LMS provides an \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emitNode} method for
programmers to control what is generated for each kind of node. Here we have
matched \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} and will generate definitions for them.
For compiled closures, we put the function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} and its accompanying syntactic term
and environment to its next-stage value representation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompiledClo}. For
applications, we have the knowledge that \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fun} is a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompiledClo} at the stage,
and it has a callable field \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, so we generate a next-stage function
application \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f(arg, σ)}. The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{quote}s are invocations to generate code for the
other next-stage values.
\begin{lstlisting}
override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
  case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar${quote(f)}, $\textdollar${quote(λ)}, $\textdollar${quote(ρ)})")
  case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar${quote(rand)}, $\textdollar${quote(σ)})")
  ...
}
\end{lstlisting}

>>>>

The IR nodes are manipulated by the LMS passes and finally generated to
next-stage Scala code. To generate code, LMS provides an \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emitNode} method for
programmers to control what is generated for each kind of node. Here we have
matched \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} and will generate definitions via @emitValDef@ for them.
For compiled closures, we put the function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} and its accompanying syntactic term
and environment to its next-stage value representation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompiledClo}. For
applications, we have the knowledge that \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fun} is a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompiledClo} at the stage,
and it has a callable field \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, so we generate a next-stage function
application \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f(arg, σ)}. The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{quote}s are invocations to generate code for the
other next-stage values.
\begin{lstlisting}
override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
  case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar${quote(f)}, $\textdollar${quote(λ)}, $\textdollar${quote(ρ)})")
  case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar${quote(rand)}, $\textdollar${quote(σ)})")
  ...
}
\end{lstlisting}

### 2019-08-16 paper/60_cases.tex
update

We borrow the notation traditional modular analysis
\cite{DBLP:conf/cc/CousotC02} and realize that such generated function is a
reusable artifact which can be used modularly.  The generated functions already
eliminate the interpretation overhead and monadic overhead, therefore can be
consider as partial summaries. Although, we still need to run the generated
function at next stage to obtain the analysis result.  At next stage, we can
provide different abstract arguments, store and caches, depending on the
context where it is called, thus the analysis is modular.

Compared with traditional summary-based analysis, our partial
summaries are represented by executable functions that are
parameterized to the context information. In our approach, the
transition from a whole-program analyzer to modular staged analyzer is
mechanized and for free from staging. We do not need to design any
intermediate representation or data structures of summaries.  In some
scenario, traditional summary-based analyses are able to carry more
information within a module, e.g., the summary of their internal
sub-modules. The staged approach presented here may still need to run
through the fixed-point for sub-modules.

>>>>

We borrow the notation traditional modular analysis
\cite{DBLP:conf/cc/CousotC02} and realize that such generated function is a
reusable artifact which can be used modularly, i.e., independent to its context. 
The generated functions already eliminate the interpretation overhead and
monadic overhead, therefore can be consider as partial summaries. Although, we
still need to run the generated function at next stage to obtain the analysis
result.  At next stage, we can provide different abstract arguments, store and
caches, depending on the context where it is called.

Compared with traditional summary-based analysis, our partial
summaries are represented by executable functions that are
parameterized to the context information. In our approach, the
transition from a whole-program analyzer to modular staged analyzer is
mechanized and for free via staging. We do not need to design any
intermediate representation or data structures of summaries.  In some
scenario, traditional summary-based analyses are able to carry more
information within a module, e.g., the summary of their internal
sub-modules. The staged approach presented here may still need to run
through the fixed-point for sub-modules.

### 2019-08-16 paper/70_eval.tex
update

Figure~\ref{evaluation_result} shows the evaluation result, comparing
the performance improved by staging on the two monovariant CFAs 
The \textit{unstaged} and \textit{staged} columns show the median time to
finish the analysis. The column $\frac{\text{unstaged}}{\text{staged}}$ shows
the improvement from staging. The dash - in some cells indicates timeout.  As
we can see from the table, the staged version significantly outperforms the
unstaged version on all benchmarks.
For 0CFA w/o store-widening, we observe an average speed-up of ~17x times.  For
0CFA w/ store-widening, the average speed-up is ~7x times.  The @kcfa-worst-64@
benchmark has the most speed-up among all the tests. Although it has a
comparablely large size of AST, it has a quite simple structure which only
involves booleans, functions and applications. We conjecture that the simple
domain and structure allow the generated code to be efficiently optimized by
the JVM. But on the corresponding 0CFA-SW, we do not observe such great
speed-up. Due to the complexity of generated code and next-stage running
platform (JVM), we leave the detailed empirical analysis of which static
analysis can benefit more from specialization as a future work.

>>>>

Figure~\ref{evaluation_result} shows the evaluation result, comparing the
performance improved by staging on the two monovariant CFAs The
\textit{unstaged} and \textit{staged} columns show the median time to finish
the analysis. The column $\frac{\text{unstaged}}{\text{staged}}$ shows the
improvement. The dash - in some cells indicates timeout.  As
expected, the staged version significantly outperforms the unstaged version on
all benchmarks. We observe that simple and homogeneous domain used in the
benchmark usualy leads to better speed-ups.  For 0CFA w/o store-widening, we
observe an average speed-up of ~17x times; for 0CFA w/ store-widening, the
average speed-up is ~7x times.  The @kcfa-worst-64@ benchmark has the most
speed-up among all the tests. Although it has a comparablely large size of AST,
it has a quite simple structure which only involves booleans, functions and
applications. We conjecture that the simple domain and structure allow the
generated code to be efficiently optimized by the JVM. But on the corresponding
0CFA-SW, we do not observe such great speed-up. Due to the complexity of
generated code and next-stage running platform (JVM), we leave the detailed
empirical analysis of which static analysis can benefit more from
specialization as future work.
